---
title: "Flyland"
output: html_notebook
---
install packages
```{r eval=FALSE, include=FALSE}
library(devtools)
install_github("rethomics/behavr")
install_github("rethomics/damr")
install_github("rethomics/ggetho")
install_github("rethomics/zeitgebr")
```

load packages
```{r}
library(zeitgebr)
library(ggetho)
library(sleepr)
```

Load package and workspace
```{r}
library(behavr)
setwd("C:/Users/bwillia/Desktop/Flyland-ALL")
DATA_DIR <- "C:/Users/bwillia/Desktop/Flyland-ALL"
list.files(DATA_DIR, pattern= "*.txt|*.csv")
setwd(DATA_DIR)
```
Metadata
```{r}
library(damr)
metadata <- fread("flyland_metadata.csv")
metadata
```
Linking
```{r}
metadata <- link_dam_metadata(metadata, result_dir = DATA_DIR)
metadata
```
Loading BASELINE flies
```{r eval=FALSE, include=FALSE}
FL_baseline <- load_dam(metadata[status == "BASELINE"])
```


Loading FL3
```{r}
#FLM_dt <- load_dam(metadata[status == "OK" & sex == "M" & generation == "FL3" & Date %in% c("180718", "180724", "180725")], FUN = sleepr::sleep_dam_annotation)

#FLF_dt <- load_dam(metadata[status == "OK"& sex == "F"& generation == "FL3" & Date %in% c("180718", "180724", "180725")], FUN = sleepr::sleep_dam_annotation)

#saveRDS(FLM_dt, file = "FL3_M_dt.RDS")
#saveRDS(FLF_dt, file = "FL3_F_dt.RDS")

FL3M_dt <- readRDS("FL3_M_dt.RDS")
FL3F_dt <- readRDS("FL3_F_dt.RDS")
```
Loading FL4
```{r}
#FLM_dt <- load_dam(metadata[status == "OK" & sex == "M" & generation == "FL4" & Date %in% c("180717", "180718","180724","180731","180801","180807", "180808")], FUN = sleepr::sleep_dam_annotation)
#FLF_dt <- load_dam(metadata[status == "OK"& sex == "F"& generation == "FL4" & Date %in% c("180717", "180718","180724","180731","180801","180807", "180808")], FUN = sleepr::sleep_dam_annotation)

#saveRDS(FLM_dt, file = "FL4_M_dt.RDS")
#saveRDS(FLF_dt, file = "FL4_F_dt.RDS")

FL4M_dt <- readRDS("FL4_M_dt.RDS")
FL4F_dt <- readRDS("FL4_F_dt.RDS")
```
Loading FL5
```{r}
#FLM_dt <- load_dam(metadata[status == "OK" & sex == "M" & generation == "FL5" & Date %in% c("180814", "180815", "180821", "180822")], FUN = sleepr::sleep_dam_annotation)
#FLF_dt <- load_dam(metadata[status == "OK" & sex == "F" & generation == "FL5" & Date %in% c("180814", "180815", "180821", "180822")], FUN = sleepr::sleep_dam_annotation)

#saveRDS(FLM_dt, file = "FL5_M_dt.RDS")
#saveRDS(FLF_dt, file = "FL5_F_dt.RDS")

FL5M_dt <- readRDS("FL5_M_dt.RDS")
FL5F_dt <- readRDS("FL5_F_dt.RDS")
```
Loading FL6
```{r}
#FLM_dt <- load_dam(metadata[status == "OK" & sex == "M" & generation == "FL6" & Date %in% c("180821", "180822", "180828", "180829")], FUN = sleepr::sleep_dam_annotation)
#FLF_dt <- load_dam(metadata[status == "OK" & sex == "F" & generation == "FL6" & Date %in% c("180821", "180822", "180828", "180829")], FUN = sleepr::sleep_dam_annotation)

#saveRDS(FLM_dt, file = "FL6_M_dt.RDS")
#saveRDS(FLF_dt, file = "FL6_F_dt.RDS")

FL6M_dt <- readRDS("FL6_M_dt.RDS")
FL6F_dt <- readRDS("FL6_F_dt.RDS")
```



UID
```{r}
#FL_baseline[, uid := paste(Monitor, Date, region_id, sep = "."), meta=T]
#FL_baseline[, .(id, uid) ,meta=T]

FL6M_dt[, uid := paste(Monitor, Date, region_id, sep = "."), meta=T]
FL6M_dt[, .(id, uid) ,meta=T]
FL6F_dt[, uid := paste(Monitor, Date, region_id, sep = "."), meta=T]
FL6F_dt[, .(id, uid) ,meta=T]
```

```{r}
rejoin(FL6M_dt)
rejoin(FL6F_dt)

FL6M_dt
FL6F_dt
```
```{r eval=FALSE, include=FALSE}
rejoin(FL_baseline)
FL_baseline
```

Remove dead animals
```{r}
library(sleepr)
FLM_curated <- curate_dead_animals(FL3M_dt)
summary(FLM_curated)
FLF_curated <- curate_dead_animals(FL3F_dt)
summary(FLF_curated)
```
See who is ded
```{r}
setdiff(FL3M_dt[, id, meta=T],
        FLM_curated[, id, meta=T])
setdiff(FL3F_dt[, id, meta=T],
        FLF_curated[, id, meta=T])
```

```{r eval=FALSE, include=FALSE}
LifespanM <- FLM_curated[xmv(drug_condit) == "5AMPH"]


LifespanF <- FLF_curated[xmv(drug_condit) == "5AMPH"]

```

```{r eval=FALSE, include=FALSE}
library(ggetho)
LifeM <- ggetho(LifespanM, aes(x= t, y= id, z=moving)) +
      stat_tile_etho()
      
LifeM
ggsave("FL4M.png", plot = LifeM, device = "png", width = 30, height = 300, units = "cm", limitsize = FALSE)


LifeF<- ggetho(LifespanF, aes(x= t, y= id, z=moving)) +
      stat_tile_etho()
      
LifeF
ggsave("FL4F.png", plot = LifeF, device = "png", width = 30, height = 300, units = "cm", limitsize = FALSE)


```


FL Baseline flies
```{r}
#FL_baseline_bin <- bin_apply_all(FL_baseline,
#                    activity,
#                    x = t,
#                    x_bin_length = mins(60),
#                    FUN = sum)
#FL_baseline_binned <- rejoin(FL_baseline_bin)
```

FL3 bin over 60 min
```{r}
FL3M_dt_bin <- rejoin(bin_apply_all(FL3M_dt,
                  activity,
                  x = t,
                  x_bin_length = mins(60),
                  FUN = sum))

FL3F_dt_bin <- rejoin(bin_apply_all(FL3F_dt,
                  activity,
                  x = t,
                  x_bin_length = mins(60),
                  FUN = sum))
```

FL4 bin over 60 min
```{r}
FL4M_dt_bin <- rejoin(bin_apply_all(FL4M_dt,
                  activity,
                  x = t,
                  x_bin_length = mins(60),
                  FUN = sum))

FL4F_dt_bin <- rejoin(bin_apply_all(FL4F_dt,
                  activity,
                  x = t,
                  x_bin_length = mins(60),
                  FUN = sum))
```

FL5 bin over 60 min
```{r}
FL5M_dt_bin <- rejoin(bin_apply_all(FL5M_dt,
                  activity,
                  x = t,
                  x_bin_length = mins(60),
                  FUN = sum))

FL5F_dt_bin <- rejoin(bin_apply_all(FL5F_dt,
                  activity,
                  x = t,
                  x_bin_length = mins(60),
                  FUN = sum))
```

FL6 bin over 60 min
```{r}
FL6M_dt_bin <- rejoin(bin_apply_all(FL6M_dt,
                  activity,
                  x = t,
                  x_bin_length = mins(60),
                  FUN = sum))

FL6F_dt_bin <- rejoin(bin_apply_all(FL6F_dt,
                  activity,
                  x = t,
                  x_bin_length = mins(60),
                  FUN = sum))

```


Male Activity Plots
```{r}
library(ggplot2)

mean.act.FL6M <- FL6M_dt_bin[, .(mean_act = mean(activity)),
                                # = std.error(activity)),
                                by = .(uid, drug_condit, t)]


mean.act.FL6M <- mean.act.FL6M[mean.act.FL6M$uid %in% c(
"20.180821.16",
"35.180822.13",
"13.180828.16"

)]
                  

# plot
actplot_TopFL6M<- ggplot(mean.act.FL6M, aes(x= t)) + 
  geom_line(aes(y= mean_act, col= drug_condit)) + 
  labs(title="Top 10% Sample", 
       subtitle="FL6 Males", 
       caption="Mean activity per 12 hrs",
       y = "Mean Activity",
       color=NULL) +  # title and caption
  scale_color_manual(labels = c("0AMPH", "5AMPH"), 
                     values = c("0AMPH"="#4169e1", 
                                "5AMPH"="#00ba38"))+ # line color
  scale_x_hours(name = "Time", breaks = waiver(),
  minor_breaks = waiver(), labels = waiver(), limits = NULL,
  expand = waiver(), oob = scales::censor, na.value = NA_real_,
  position = "bottom", time_wrap = NULL, unit = "h", log = FALSE)+ # set x axis scale
  
  scale_y_continuous(limits = c(0, 1500)) + # set y axis scale
  
  #geom_errorbar(aes(ymin=mean_act-sem, ymax=mean_act+sem), width=.2,
                # position=position_dodge(0.05))+ # error bars
  facet_grid(rows = vars(uid), cols = NULL, scales = "fixed",
  space = "fixed", shrink = TRUE, labeller = "label_value",
  as.table = TRUE, switch = NULL, drop = TRUE, margins = FALSE,
  facets = NULL)+

  theme(axis.text.x = element_text(angle = 45, vjust=0.5, size = 8)) # rotate x axis text
    

ggsave("TopFL6M_actplot.png", plot = actplot_TopFL6M, device = "png", width = 11 , height = 7, units = "in")

actplot_TopFL6M
```
Female Activity Plots
```{r}
library(ggplot2)

mean.act.FL6F <- FL6F_dt_bin[, .(mean_act = mean(activity)),
                                # = std.error(activity)),
                                by = .(uid, drug_condit, t)]


mean.act.FL6F <- mean.act.FL6F[mean.act.FL6F$uid %in% c(
"17.180821.23",
"29.180822.17",
"6.180828.27",
"22.180829.9"

  )]


# plot
actplot_TopFL6F<- ggplot(mean.act.FL6F, aes(x= t)) + 
  geom_line(aes(y= mean_act, col= drug_condit)) + 
  labs(title="Top 10% Sample", 
       subtitle="FL6 Females", 
       caption="Mean activity per 12 hrs",
       y = "Mean Activity",
       color=NULL) +  # title and caption
  scale_color_manual(labels = c("0AMPH", "5AMPH"), 
                     values = c("0AMPH"="#4169e1", 
                                "5AMPH"="#00ba38"))+ # line color
  scale_x_hours(name = "Time", breaks = waiver(),
  minor_breaks = waiver(), labels = waiver(), limits = NULL,
  expand = waiver(), oob = scales::censor, na.value = NA_real_,
  position = "bottom", time_wrap = NULL, unit = "h", log = FALSE)+ # set x axis scale
  
  scale_y_continuous(limits = c(0, 1500)) + # set y axis scale
  
  #geom_errorbar(aes(ymin=mean_act-sem, ymax=mean_act+sem), width=.2,
                # position=position_dodge(0.05))+ # error bars
  facet_grid(rows = vars(uid), cols = NULL, scales = "fixed",
  space = "fixed", shrink = TRUE, labeller = "label_value",
  as.table = TRUE, switch = NULL, drop = TRUE, margins = FALSE,
  facets = NULL)+

  theme(axis.text.x = element_text(angle = 45, vjust=0.5, size = 8)) # rotate x axis text
    

ggsave("TopFL6F_actplot.png", plot = actplot_TopFL6F, device = "png", width = 11 , height = 11, units = "in")

actplot_TopFL6F
```

```{r}
N1_avg <- FL_baseline_binned[t >= hours(7) & t <= hours(18), .(N1_avg = mean(activity)),
                      by = .(drug_condit, uid)
                      ]


N1M_avg <- FLM_dt_binned[t >= hours(7) & t <= hours(18), .(N1_avg = mean(activity)),
                      by = .(drug_condit, uid)
                      ]



N1F_avg <- FLF_dt_binned[t >= hours(7) & t <= hours(18), .(N1_avg = mean(activity)),
                      by = .(drug_condit, uid)
                      ]

```

```{r}
FL_N1<- N1_avg[, Delta5 := N1_avg[drug_condit == "5AMPH"] - N1_avg, .(uid)][, Ratio5 := N1_avg[drug_condit == "5AMPH"] / N1_avg, .(uid)]

FLM_N1<- N1M_avg[, Delta5 := N1_avg[drug_condit == "5AMPH"] - N1_avg, .(uid)][, Ratio5 := N1_avg[drug_condit == "5AMPH"] / N1_avg, .(uid)]

FLF_N1<- N1F_avg[, Delta5 := N1_avg[drug_condit == "5AMPH"] - N1_avg, .(uid)][, Ratio5 := N1_avg[drug_condit == "5AMPH"] / N1_avg, .(uid)]
```

```{r}
N1_0AMPH <- FL_N1[drug_condit == "0AMPH"]
N1_5AMPH <- FL_N1[drug_condit == "5AMPH"]

write.csv(N1_0AMPH, file = "N1_baseline.csv")
write.csv(N1_5AMPH, file = "N1_baseline5.csv")


N1_0AMPH_M <- FLM_N1[drug_condit == "0AMPH"]
N1_5AMPH_M <- FLM_N1[drug_condit == "5AMPH"]

N1_0AMPH_F <- FLF_N1[drug_condit == "0AMPH"]
N1_5AMPH_F <- FLF_N1[drug_condit == "5AMPH"]

```
FL3
```{r}
write.csv(N1_0AMPH_M, file = "N1_0AMPH_M_FL3.csv")
write.csv(N1_5AMPH_M, file = "N1_5AMPH_M_FL3.csv")

write.csv(N1_0AMPH_F, file = "N1_0AMPH_F_FL3.csv")
write.csv(N1_5AMPH_F, file = "N1_5AMPH_F_FL3.csv")

```

FL4
```{r}
write.csv(N1_0AMPH_M, file = "N1_0AMPH_M_FL4.csv")
write.csv(N1_5AMPH_M, file = "N1_5AMPH_M_FL4.csv")

write.csv(N1_0AMPH_F, file = "N1_0AMPH_F_FL4.csv")
write.csv(N1_5AMPH_F, file = "N1_5AMPH_F_FL4.csv")
```

FL5
```{r}
write.csv(N1_0AMPH_M, file = "N1_0AMPH_M_FL5.csv")
write.csv(N1_5AMPH_M, file = "N1_5AMPH_M_FL5.csv")

write.csv(N1_0AMPH_F, file = "N1_0AMPH_F_FL5.csv")
write.csv(N1_5AMPH_F, file = "N1_5AMPH_F_FL5.csv")
```

FL6
```{r}
write.csv(N1_0AMPH_M, file = "N1_0AMPH_M_FL6.csv")
write.csv(N1_5AMPH_M, file = "N1_5AMPH_M_FL6.csv")

write.csv(N1_0AMPH_F, file = "N1_0AMPH_F_FL6.csv")
write.csv(N1_5AMPH_F, file = "N1_5AMPH_F_FL6.csv")
```














Check lifespan
```{r}
# we make a summary table of all lifespan for each animals
Mlifespan_dt <- FLM_curated[drug_condit == "5AMPH", .(lifespan = max(t)), by= id]
Flifespan_dt <- FLF_curated[drug_condit == "5AMPH", .(lifespan = max(t)), by= id]
```
Remove flies who died within 68 hrs (3days)
```{r}
# we filter this table for lifespan>2 and we keep the id
Mvalid_ids <- Mlifespan_dt[lifespan > hours(68), id]
Fvalid_ids <- Flifespan_dt[lifespan > hours(68), id]

# we apply this filter
FLM_curated2 <- FLM_curated[id %in% Mvalid_ids]
FLF_curated2 <- FLF_curated[id %in% Fvalid_ids]
```


```{r}
setdiff(FLM_curated[, id, meta=T],
        FLM_curated2[, id, meta=T])
setdiff(FLF_curated[, id, meta=T],
        FLF_dt_curated2[, id, meta=T])
```





```{r}
FLM_N1<- N1M_MeanAct[, Delta5 := N1MeanAct[drug_condit == "5AMPH"] - N1MeanAct, .(uid)][, Ratio5 := N1MeanAct[drug_condit == "5AMPH"] / N1MeanAct, .(uid)]

FLF_N1<- N1F_MeanAct[, Delta5 := N1MeanAct[drug_condit == "5AMPH"] - N1MeanAct, .(uid)][, Ratio5 := N1MeanAct[drug_condit == "5AMPH"] / N1MeanAct, .(uid)]
```

```{r}
N1_0AMPH_M <- FLM_N1[drug_condit == "0AMPH"]
N1_5AMPH_M <- FLM_N1[drug_condit == "5AMPH"]



write.csv(N1_0AMPH_M, file = "N1_0AMPH_M.csv")
write.csv(N1_5AMPH_M, file = "N1_5AMPH_M.csv")



N1_0AMPH_F <- FLF_N1[drug_condit == "0AMPH"]
N1_5AMPH_F <- FLF_N1[drug_condit == "5AMPH"]



write.csv(N1_0AMPH_F, file = "N1_0AMPH_F.csv")
write.csv(N1_5AMPH_F, file = "N1_5AMPH_F.csv")

```





```{r}
library(zeitgebr)
FLM_dt_bin[, t := ifelse(xmv(drug_condit) == "0AMPH", t, t * 1.01),]

per_dt <- periodogram(activity, FLM_dt_bin, FUN = chi_sq_periodogram, resample_rate = 1/mins(5))
ggperio(per_dt, aes(period, power, colour=drug_condit)) + 
        stat_pop_etho()
```





Check lifespan
```{r}
#FLM_curated[, lifespan := max(t), by= id]
#FLM_curated <- FLMcurated[, .(id, lifespan), meta=T]
#FLF_curated[, lifespan := max(t), by= id]
#FLF_curated[, .(id, lifespan), meta=T]

#rejoin(FLM_curated)
#rejoin(FLF_curated)
```
Subset 4day flies
```{r}
#FourDayFlies <- FLM_curated[xmv(Date) == c("180717", "180724", "180731", "181807", "180814", "180821", "180828", "180718", "180725")]
#FourDayFlies_valid <- FourDayFlies[lifespan > hours(92), id]
#FLM_curated2 <- FourDayFlies[id %in% FourDayFlies_valid]

```
See who was removed
```{r}
#setdiff(FourDayFlies[, id, meta=T],
       # FLM_curated2[, id, meta=T])

```

**TRY THIS**
```{r}
# we make a summary table of all lifespan for each animals
Mlifespan_dt <- FLM_curated[, .(lifespan = max(t)), by= id]
Flifespan_dt <- FLF_curated[, .(lifespan = max(t)), by= id]


MMetaLife <- FLM_curated[Mlifespan_dt, meta=T]
setmeta(FLM_curated, MMetaLife)

FMetaLife <- FLF_curated[Flifespan_dt, meta=T]
setmeta(FLF_curated, FMetaLife)
```
Remove flies who died within 68 hrs (3days)
```{r}
# we filter this table for lifespan>2 and we keep the id
Mvalid_ids <- Mlifespan_dt[lifespan > hours(68), id]
Fvalid_ids <- Flifespan_dt[lifespan > hours(68), id]

# we apply this filter
FLM_curated2 <- FLM_curated[id %in% Mvalid_ids]
FLF_curated2 <- FLF_curated[id %in% Fvalid_ids]


```
See who was removed
```{r}
setdiff(FLM_curated[, id, meta=T],
        FLM_curated2[, id, meta=T])
setdiff(FLF_curated[, id, meta=T],
        FLF_dt_curated2[, id, meta=T])
```

```{r}
##ThreeDayFliesM <- FLM_curated[xmv(Date) == c("181801", "180808", "180815", "180822", "180729")]

##FourDayFliesF <- FLF_curated[xmv(Date) == c("180717", "180724", "180731", "181807", "180814", "180821", "180828", "180718", "180725")]
##ThreeDayFliesF <- FLF_curated[xmv(Date) == c("181801", "180808", "180815", "180822", "180729")]
```

```{r}
###N3M_avg <- FLM_dt_binned[t %between% c(hours(56), hours(67)), 
###                       .(N3M_avg = mean(activity)),
###                        by = .(drug_condit, uid)
###                        ]
###N3M_MeanAct <- N3M_avg[, .("N3MeanAct" = mean(activity)),
###                 by = .(drug_condit, uid)]

###N3F_avg <- FLF_dt_binned[t %between% c(hours(56), hours(67)), 
###                       .(N3F_avg = mean(activity)),
###                        by = .(drug_condit, uid)
###                        ]
###N3F_MeanAct <- N3F_avg[, .("N3MeanAct" = mean(activity)),
###                 by = .(drug_condit, uid)]

```